# Loading in Libraries

```{r}
#Loading in Libraries

library(tidyverse)
library(ggthemes)
library(readxl)
library(lubridate)
library(purrr)
library(moments) #used to calculate skewness and kurtosis

l_branches <- c("lib1", "lib2")
m_branches <- c("lib3", "lib4")
```


# Circulation Code  

## daily circulation statistics 

### Data cleaning for daily circulation dataset "circ_day"
```{r}
orig_circ_day <- read_xlsx("Data/Example Data.xlsx", sheet = "phys_circ_day")

orig_circ_day$date<- as.Date(orig_circ_day$date)

#pivot to wider graph
circ_day <- pivot_wider(orig_circ_day, names_from = branch, values_from = transactions)

#Creating a synthetic month variable 
#circ$date <- as.Date(paste(circ$year, circ$month, "01", sep = "-"))

#Group circulation by week to better display that on a graph 
circ_day$Week_Num <- ifelse(circ_day$day == "Monday", 1, 0)

#Sorting the dataframe by date 
circ_day <- arrange(circ_day, date)

#Properly counting the week number 
circ_day$week_number <- cumsum(circ_day$Week_Num)

#However before I remove missing values I need to create a data frame averaged by week
week_dates <- circ_day %>% filter(Week_Num == 1) %>% summarise(date, week_number) 

#Cleaning the data to remove any rows that have a missing value, idk if I'll keep this or not 
#circ_day <- as.data.frame(circ_day[complete.cases(circ_day),])

#Converting all relevant columns to numeric variables 
only_nums <- circ_day %>% select(-date, -day)
only_dates <- circ_day %>% select(date, day)

only_nums <- as.data.frame(lapply(only_nums, as.numeric))
circ_day <- cbind(only_dates, only_nums)

class(circ_day$lib1)

#Cleaning
rm(only_nums, only_dates)
```


```{r}
#Checking to make sure there's no more than 7 days in each week 
circ_day %>% group_by(week_number) %>% summarise(n = n())

#Other ways of extracting week number, except these dates reset every year 

# extract the week number using %U
#   circ_day$week_num_U2 <- strftime(circ_day$date, format = "%U")


# extract the week number using %W
#the %W code returns the week number of the year with Monday as the first day of the week.
#   circ_day$week_num_U2 <- strftime(circ_day$date, format = "%W")
```


### Examing checkouts by day at an individual branch (over time and boxplot by day)
```{r}
#Choosing branch to analyze for daily checkouts

#Branches: large: "lib1", "lib2" medium "lib3", "lib4"
branch_circ_day <- "lib1"
```


```{r}
ggplot(circ_day, aes(date, .data[[branch_circ_day]])) +
  geom_line(size = 1, color = "blue3") +
  geom_point(size = 2) +
  
  labs(
    title = paste0("Number of Checkouts per Day for ", branch_circ_day, " in Q1 2023"),
    y = "Checkouts",
    x = NULL,
    caption = "Figure 2.1") + 
  
  scale_y_continuous(
    breaks = seq(0, 1200, by = 100),
    limits = c(0, 1200)) +
  
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%B %d",
    limits = as.Date(c("2023-04-01", "2023-07-31"), "%Y-%m-%d"),
        expand = c(0.01, 0)) +
  
  theme_clean() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        #Adding a vertical line at each week interval 
        panel.grid.major.x = element_line(color = "grey", size = 0.2))
```


```{r}
#Examing distribution of checkouts by day at an individual branch using multiple boxplots 

#Need to change the day variable to a factor variable to specify the order so that it appears nicely on the graph 
circ_day$day <- factor(circ_day$day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

#filtering the time we want to looks at (Q1 2023)
start_date_box <- "2023-04-01"
end_date_box <- "2023-6-30"
circ_day_daily <- circ_day %>% filter(date >= as.Date(start_date_box, "%Y-%m-%d") & date < as.Date(end_date_box, "%Y-%m-%d"))

#The graph
ggplot(circ_day_daily, aes(day, .data[[branch_circ_day]])) +
  
  geom_boxplot(fill = "blue1",  color = "black", outlier.color = "red") +
  
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2, fill = "white") + #adding the mean to the boxplots
  
  labs(title = paste0("Transactions by Day at ",  branch_circ_day),
       subtitle = paste0("From ", format(as.Date(start_date_box), "%B %d, %Y"), " to ", format(as.Date(end_date_box), "%B %d, %Y")),
       x = NULL,
       y = "Checkouts",
       caption = "Figure 1.7") + 
  
  scale_y_continuous(
    breaks = seq(0, 1200, by = 100),
    limits = c(0, 1200)) +
  
  #scale_x_discrete(labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  
  #scale_fill_manual(values = c("grey85", "#FFDDCC"), guide = FALSE) +  #could potentially be used to highlight one of the boxplots 
  
  theme_clean() +
  theme(panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
        plot.subtitle = element_text(size = 10),
        plot.caption = element_text(size = 8))
```



```{r}
# Distribution plot example - The geom_density() function in R is used to create a density plot, which visualizes the distribution of a continuous variable. It estimates the underlying probability density function (PDF) based on the data and displays it as a smooth curve.

#Could do an intersitng thing with this graph 

median <- median(circ_day$lib1, na.rm = TRUE)
mean <- round(mean(circ_day$lib1, na.rm = TRUE), digits = 0)

ggplot(circ_day, aes(lib1)) + 
  geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   colour="black", fill="white")+
  geom_density(alpha = .2, fill = "pink2") + 
  
  labs(title = "Checkout Distribution at Lib1",
       x = "Checkouts",
       y = "Density") +
  
  geom_vline(xintercept = median, color = "red") +
  geom_text(aes(x = 1000, y = .0015, label = paste0("Median: ", median)), 
            vjust = 0, hjust = .4, size = 4) +  # fontface = "bold.italic")
  geom_curve(x = 1000, y = .0015, xend = (median + 10), yend = .0010, 
             curvature = -.3, arrow = arrow(length = unit(2, "mm"))) +
  
  geom_vline(xintercept = mean, color = "blue") +
  geom_text(aes(x = 370, y = .0015, label = paste0("Mean: ", mean)), 
            color = "blue", vjust = 0, hjust = .4, size = 4) +
  geom_curve(x = 370, y = .0015, xend = (mean - 10), yend = .0010, 
             curvature = .3, arrow = arrow(length = unit(2, "mm"))) +
  
  geom_text(aes(x = -Inf, y = Inf, label = paste0("Kurtosis: ", round(kurtosis(circ_day$lib1, na.rm = TRUE), digits = 2), "\n", "Skewness: ", round(skewness(circ_day$lib1, na.rm = TRUE), digits = 2)), hjust = -0.05, vjust = 1.1)) +

  theme_clean() +
    
  theme(plot.title = element_text(hjust = .5, color = "purple", size = 16, face = "bold"))

rm(median, mean)
```



### Examing checkouts by day at various large branches (boxplots)


#### Data wrangling
```{r}
#filtering the time we want to looks at (Q1 2023)
start_date <- "2023-04-01"
end_date <- "2023-6-30"

boxplot_circ_day <- orig_circ_day %>% filter(date >= as.Date(start_date, "%Y-%m-%d") & date < as.Date(end_date, "%Y-%m-%d"))

#Reordering the dataset to display by descending order in the box plot

group_ordered <- with(boxplot_circ_day, reorder(branch, desc(transactions), median))

boxplot_circ_day$branch <- factor(boxplot_circ_day$branch, levels = levels(group_ordered))

# Assigning colors based on median value
boxplot_circ_day$branch_size <- ifelse(boxplot_circ_day$branch %in% l_branches, "l_branch",
                                       ifelse(boxplot_circ_day$branch %in% m_branches, "m_branch", "s_branch"))
```


#### Summary statistics for circulation per day at every branch using boxplots
```{r}
#For this graph need to watch out how saturday and sunday affect distribution because of different branch hours

ggplot(boxplot_circ_day, aes(branch, transactions, fill = branch_size)) +
  
  geom_boxplot(color = "black", outlier.color = "red") +
  
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2, fill = "white") + #adding the mean to the boxplots
  
  labs(title = paste0("Summary Statistics for Circulation by Day"),
      subtitle = paste0("From ", format(as.Date(start_date), "%B %d, %Y"), " to ", format(as.Date(end_date), "%B %d, %Y")), 
       x = NULL,
       y = "Checkouts",
       fill = NULL) +

  scale_y_continuous(
    breaks = seq(0, 2250, by = 500),
    limits = c(0, 2250)) +
  
  scale_fill_manual(values = c("l_branch" = "#f0645d", "m_branch" = "#fcb813", "s_branch" = "#00b059"), ##1e196a
                    labels = c("l_branch" = "Large Branch", "m_branch" = "Medium Branch", "s_branch" = "Small Branch")) +
  
  theme_clean() +
  theme(panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
        axis.text.x = element_text(angle = 25, hjust = 1),
        plot.subtitle = element_text(size = 18),
        plot.title = element_text(size = 25),
        legend.position = c(0.87, 0.81), 
        legend.text = element_text(size = 14))
```



## Examing checkouts by week at large and medium branches 

### Data Wrangling
```{r}
circ_week_list <- list()

for (branch in l_branches){
circ_day_week <- circ_day %>% 
  group_by(week_number) %>% 
  summarise(avg_week = mean(!!sym(branch), na.rm = TRUE)) %>%
  rename(!!paste0(branch, "_week_avg") := avg_week)
  
  circ_week_list[[branch]] <- circ_day_week 
  
  rm(circ_day_week)
}

#This code creates a new data frame circ_day_week_combined that has columns for each branch in l_branches, with the suffix _week_avg added to the column name. The across() function is used to apply the mean() function to each column specified in all_of(l_branches). The .names argument in across() is used to generate the new column names by using the {col} placeholder to insert the original column name.

circ_week_combined <- circ_day %>% 
  group_by(week_number) %>% 
  summarise(across(all_of(m_branches), ~mean(.x, na.rm = TRUE), .names = "{col}_week_avg")) #l_branches or m_branches

circ_week <- merge(week_dates, circ_week_combined, by = "week_number")

rm(circ_week_combined, circ_week_list)
```


### Graphing checkouts by week at large branches 
```{r eval=FALSE, include=FALSE}
ggplot(circ_week, aes(date)) +
  geom_line(aes(y = lib1_week_avg, color = "lib1"), size = 1) +
  geom_point(aes(y = lib1_week_avg)) +
  
  geom_line(aes(y = lib2_week_avg, color = "lib2"), size = 1) +
  geom_point(aes(y = lib2_week_avg)) +
  
  labs(
    title = "Average Daily checkout per Week Across Large Branches",
    y = "Checkouts",
    x = NULL,
    caption = "Figure 2.3",
    subtitle = "Q2 2023") + 
  
  scale_y_continuous(
    breaks = seq(0, 1200, by = 100),
    limits = c(0, 1200)) +
  
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%B %d",
    limits = as.Date(c("2023-04-01", "2023-06-30"), "%Y-%m-%d"), 
    expand = c(.002, 0)) +
  
   scale_color_manual(name = NULL,
                     values = c("lib1" = "blue3", "lib2" = "black")) +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 20, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      plot.subtitle = element_text(size = 12),
      legend.position = c(0.4, 0.88),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8)) 
```


### Graphing checkouts by week at medium branches 
```{r}
ggplot(circ_week, aes(date)) +

  geom_line(aes(y = lib3_week_avg, color = "Library 3"), linewidth = 1) +
  geom_point(aes(y = lib3_week_avg)) +
  
  geom_line(aes(y = lib4_week_avg, color = "Library 4"), linewidth = 1) +
  geom_point(aes(y = lib4_week_avg)) +

  labs(
    title = "Average Daily Checkout Across Branches",
    y = "Checkouts",
    x = NULL,
    caption = "Figure 2.4",
    subtitle = "Q2 2023") + 
  
  scale_y_continuous(
    breaks = seq(0, 1200, by = 100),
    limits = c(0, 1200)) +
  
  scale_x_date(
    date_breaks = "1 week",
    date_labels = "%B %d",
    limits = as.Date(c("2023-04-01", "2023-06-30"), "%Y-%m-%d"), 
    expand = c(.002, 0)) +
  
   scale_color_manual(name = NULL,
                     values = c("Library 3" = "hotpink", "Library 4" = "green")) +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 20, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      plot.subtitle = element_text(size = 12),
      legend.position = c(0.4, 0.88),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8)) 
```




## monthly circulation statistics 


### Data cleaning for monthly circulation dataset "circ_mon"

```{r}
#Data cleaning 

circ_mon <- read_xlsx("Data/Example Data.xlsx", sheet = "phys_circ_mon")

#pivot to wider graph
circ_mon <- pivot_wider(circ_mon, names_from = branch, values_from = transactions)

#Creating a synthetic date variable 
circ_mon$date <- as.Date(paste(circ_mon$year, circ_mon$month, "01", sep = "-"))

class(circ_mon$date)

#grabbing the month and year (12-18 for december 2018) of the tweet tweeted for later analysis to group by month and year 
circ_mon$month <- format(circ_mon$date, "%b") 
```


### Circulation numbers by month for 1 branch (linegraphs) 
```{r}
#Choosing a branch to examine for monthly checkouts 

#Branches: large: "lib1", "lib2" medium "lib3", "lib4"

branch_circ_mon <- "lib2"
```


```{r}
ggplot(circ_mon, aes(date, .data[[branch_circ_mon]])) +
  geom_point(size = 2) +
  geom_line(size = 1, color = "blue3") +
  geom_smooth(color = "black", se = FALSE) +
  
  labs(
    title = paste0(branch_circ_mon, " Checkouts 2021-2023"),
    y = "Checkouts",
    x = "Date",
    caption = "Figure 1.4"
  ) + 
  
  #scale_y_continuous(
    #breaks = seq(16000, 27500, by = 1000),
    #limits = c(16000, 27500)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2021-01-01", "2023-07-01"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    expand = c(.01, 0)) +
  
  annotate("rect", xmin = as.Date("2022-01-01", "%Y-%m-%d"), xmax = as.Date("2022-03-01", "%Y-%m-%d"), 
           ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  annotate("rect", xmin = as.Date("2023-01-01", "%Y-%m-%d"), xmax = as.Date("2023-03-01", "%Y-%m-%d"), 
           ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  annotate("rect", xmin = as.Date("2021-01-01", "%Y-%m-%d"), xmax = as.Date("2021-03-01", "%Y-%m-%d"), 
           ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"))
```


```{r}
ggplot(circ_mon, aes(month(date, label=TRUE, abbr=TRUE), .data[[branch_circ_mon]], group = factor(year(date)), color = factor(year(date)))) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  #geom_smooth(color = "black", se = FALSE) +
  
  labs(
    title = paste0("Monthly Checkouts at " , branch_circ_mon), 
    subtitle = "From 2021 - 2023",
    y = "Checkouts",
    x = NULL,
    caption = "Figure 2.2") + 
  
  #scale_y_continuous(
    #breaks = seq(0, 15000, by = 1500),
    #limits = c(0, 15000)) +
  
  scale_color_manual(name = NULL,
                     values = #c("2021" = "#f0645d", "2022" = "#fcb813", "2023" = "#00b059")) +
                              c("2021" = "chartreuse3", "2022" = "orangered2", "2023" = "blue3")) +
  theme_clean() +
  
  theme(#Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      plot.subtitle = element_text(size = 12),
      legend.position = c(0.9, 0.85), 
      legend.text = element_text(size = 11))
```


### Circulation numbers by month for large branches 
```{r}
ggplot(circ_mon, aes(date)) +

  geom_line(aes(y = lib1, color = "Library 1"), size = 1) +
  geom_point(aes(y = lib1), size = 2) + 
  
  geom_line(aes(y = lib2, color = "Library 2"), size = 1) +
  geom_point(aes(y = lib2), size = 2) + 
  
  labs(
    title = "Checkouts per Month Across Large Branches Since 2021",
    y = "Checkouts",
    x = NULL,
    caption = "Figure 2.4"
  ) + 
  
  #scale_y_continuous(
    #breaks = seq(0, 35000, by = 2500),
    #limits = c(0, 35000)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2021-01-01", "2023-06-01"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    #For some reason this needs to be added to stop ggplot from expanding the x axis despite the specification of the limit
    expand = c(.01, 0)) +
  
  scale_color_manual(name = NULL,
                     values = c("Library 1" = "blue3", "Library 2" = "black")) +
  
  #annotate("rect", xmin = as.Date("2022-01-01", "%Y-%m-%d"), xmax = as.Date("2022-03-01", "%Y-%m-%d"), 
  #         ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2023-01-01", "%Y-%m-%d"), xmax = as.Date("2023-03-01", "%Y-%m-%d"), 
  #        ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2021-01-01", "%Y-%m-%d"), xmax = as.Date("2021-03-01", "%Y-%m-%d"), 
  #         ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.84, 0.25),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8))
```


### Circulation numbers by month for medium branches 
```{r}
ggplot(circ_mon, aes(date)) +
  geom_line(aes(y = lib3, color = "Library 3"), size = 1) +
  geom_point(aes(y = lib3), size = 2) + 

  geom_line(aes(y = lib4, color = "Library 4"), size = 1) +
  geom_point(aes(y = lib4), size = 2) +
  
  labs(
    title = "Checkouts per Month Across Medium Branches Since 2021",
    y = "Checkouts",
    x = NULL,
    caption = "Figure 2.5"
  ) + 
  
  #scale_y_continuous(
    #breaks = seq(0, 30000, by = 2000),
    #limits = c(0, 30000)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2021-01-01", "2023-06-01"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    #For some reason this needs to be added to stop ggplot from expanding the x axis despite the specification of the limit
    expand = c(.01, 0)) +
  
  scale_color_manual(name = NULL,
                     values = c("Library 3" = "hotpink",  "Library 4" = "green")) +
  
  #annotate("rect", xmin = as.Date("2022-01-01", "%Y-%m-%d"), xmax = as.Date("2022-03-01", "%Y-%m-%d"), 
  #         ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2023-01-01", "%Y-%m-%d"), xmax = as.Date("2023-03-01", "%Y-%m-%d"), 
  #        ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2021-01-01", "%Y-%m-%d"), xmax = as.Date("2021-03-01", "%Y-%m-%d"), 
  #         ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.80, 0.88),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8))
```


## plotting collection size vs percentage of checkouts for each area  
```{r eval=FALSE, include=FALSE}
# No example data to run for this code

humms_long <- pivot_longer(humms, cols = c("checkouts", "collection_size"), names_to = "type", values_to = "percent")


ggplot(humms_long, aes(x = Book_type, y = percent, fill = type)) +
  geom_bar(stat = "identity", position ="dodge", color = "black") +
  
  labs(title = "Total Checkouts VS. Size of Collection", 
       subtitle = "lib Q2 2023",
       x = NULL, 
       y = "Percent (%)",
       caption = "Figure 2.5") +
  
   scale_y_continuous(
      breaks = seq(0, 25, by = 5),
      limits = c(0, 25)) +
  
  scale_fill_manual(name = NULL,
                    labels= c("Total Checkouts", "Size of Collection"),
                    values = c("checkouts" = "chartreuse3", "collection_size" = "firebrick2")) +
  
  theme_clean() +
  theme(panel.grid.minor.y = element_line(color = "gray", linetype = "dotted"),
        plot.subtitle = element_text(size = 10),
        legend.position = c(0.84, 0.80),
        legend.text = element_text(size = 10))
```


## Circulation charts (not graphs) summary and descriptive statistics  
```{r}
#Getting the summary stats for the Q1 lib1
trial <- circ_day %>% filter(date >= as.Date("2023-04-01", "%Y-%m-%d") & date < as.Date("2023-06-30", "%Y-%m-%d"))

data.frame(unclass(summary(trial$lib1)))

#Getting the summary stats for the Q1 lib2
trial <- circ_day %>% filter(date >= as.Date("2023-01-01", "%Y-%m-%d") & date < as.Date("2023-04-01", "%Y-%m-%d"))

data.frame(unclass(summary(trial$lib2)))

#Getting average number of checkouts per day over Q1 2023 
circ_day %>% filter(date >= as.Date("2023-04-01", "%Y-%m-%d") & date < as.Date("2023-06-30", "%Y-%m-%d")) %>% 
  group_by(day) %>% 
  summarise(Average_Checkouts = mean(lib1, na.rm = TRUE), n = n()) %>% 
  arrange(desc(Average_Checkouts))
```


# Examing new card registrations

## Data cleaning

```{r}
new_cards <- read_xlsx("Data/Example Data.xlsx", sheet = "new_card")

#pivot to wider graph
new_cards <- pivot_wider(new_cards, names_from = branch, values_from = new_card_accounts)

#Creating a synthetic month variable 
new_cards$date <- as.Date(paste(new_cards$year, new_cards$month, "01", sep = "-"))

#grabbing the month and year (12-18 for december 2018) of the tweet tweeted for later analysis to group by month and year 
new_cards$month <- format(new_cards$date, "%b") 
```


## New card graphs (line and bar) for Individual branch
```{r}
## Choosing the branch for new card anaylsis

#Branches: large: "lib1", "lib2" medium "lib3", "lib4"

branch_new_cards <- "lib1"
```


```{r}
#Line graph 

ggplot(new_cards, aes(date, .data[[branch_new_cards]])) + 
  geom_line() +
  geom_point(size = 2) +
  
  labs(
    title = paste0("New Card Registrations at ", branch_new_cards), 
    y = "New Cards",
    x = "Date"
  ) + 
  
  scale_y_continuous(
    breaks = seq(0, 300, by = 25),
    limits = c(0, 300)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2022-01-01", "2022-12-11"), "%Y-%m-%d"),
    date_labels = "%b '%y") +
  
  theme_clean()
```


```{r}
#bar graph with numbers on top of bars 

ggplot(new_cards, aes(date, .data[[branch_new_cards]])) + 
  geom_bar(stat = "identity", color = "black", fill = "blue3") +
  geom_label(aes(label = .data[[branch_new_cards]]), hjust = .5, nudge_y = 7.3) +
  
  labs(
    title = paste0("New Card Registrations at ", branch_new_cards, " since 2022"),
    y = "Number of New Cards",
    x = NULL,
    caption = "Figure 5.1"
  ) + 
  
  scale_y_continuous(
    breaks = seq(0, 300, by = 25),
    limits = c(0, 300)) +
  
   scale_x_date(
    date_breaks = "month",
    limits = as.Date(c("2021-12-14", "2023-06-15"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    expand = c(0.01, 0)) +

  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```


## New cards across Large Branches
```{r}
ggplot(new_cards, aes(date)) +
  geom_line(aes(y = lib1, color = "Library 1"), size = 1) +
  geom_point(aes(y = lib1), size = 2) + 

  geom_line(aes(y = lib2, color = "Library 2"), size = 1) +
  geom_point(aes(y = lib2), size = 2) + 
  
  labs(
    title = "New Card Registrations per Month Across Large Branches",
    y = "New Cards",
    x = NULL,
    caption = "Figure 5.2"
  ) + 
  
  scale_y_continuous(
    breaks = seq(0, 300, by = 50),
    limits = c(0, 300)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2021-08-01", "2023-06-01"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    #For some reason this needs to be added to stop ggplot from expanding the x axis despite the specification of the limit
    expand = c(.01, 0)) +
  
  scale_color_manual(name = NULL,
                     values = c("Library 1" = "blue3", "Library 2" = "black")) +
  
  #annotate("rect", xmin = as.Date("2022-01-01", "%Y-%m-%d"), xmax = as.Date("2022-03-01", "%Y-%m-%d"), 
  #         ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2023-01-01", "%Y-%m-%d"), xmax = as.Date("2023-03-01", "%Y-%m-%d"), 
  #        ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2021-01-01", "%Y-%m-%d"), xmax = as.Date("2021-03-01", "%Y-%m-%d"), 
  #         ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.74, 0.9),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8))
```


## New cards across Medium Branches
```{r}
ggplot(new_cards, aes(date)) +
  geom_line(aes(y = lib3, color = "Library 3"), size = 1) +
  geom_point(aes(y = lib3), size = 2) + 

  geom_line(aes(y = lib4, color = "Library 4"), size = 1) +
  geom_point(aes(y = lib4), size = 2) + 
  
  labs(
    title = "New Card Registrations per Month Across Medium Branches",
    y = "New Cards",
    x = NULL,
    caption = "Figure 5.2"
  ) + 
  
  scale_y_continuous(
    breaks = seq(0, 300, by = 50),
    limits = c(0, 300)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2022-01-01", "2023-06-01"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    #For some reason this needs to be added to stop ggplot from expanding the x axis despite the specification of the limit
    expand = c(.01, 0)) +
  
  scale_color_manual(name = NULL,
                     values = c("Library 3" = "hotpink", "Library 4" = "green")) +

  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.74, 0.84),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8))
```



## New cards across specific branch Year over year
```{r}
ggplot(new_cards, aes(month(date, label=TRUE, abbr=TRUE), group = factor(year), color = factor(year))) +
  
  geom_line(aes(y = lib3), size = 1) +
  geom_point(aes(y = lib3), size = 2) + 

  labs(
    title = "New Card Registrations per Month at lib3",
    y = "New Cards",
    x = NULL,
    caption = "Figure 5.3",
    color = NULL) + 
  
  scale_y_continuous(
    breaks = seq(0, 300, by = 50),
    limits = c(0, 300)) +
  
  scale_color_manual(name = NULL,
                     values = c("2021" = "chartreuse3", "2022" = "orangered2", "2023" = "blue3")) +
  
  theme_clean() +
  
  theme(#Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.84, 0.84),
      legend.text = element_text(size = 10))
```

## Pure Summary statistics 
```{r}
#Getting average number of checkouts per day over Q1 2023 
circ_day %>% filter(date >= as.Date("2023-04-01", "%Y-%m-%d") & date < as.Date("2023-06-30", "%Y-%m-%d")) %>% 
  group_by(day) %>% 
  summarise(Average_Checkouts = mean(lib2, na.rm = TRUE), n = n()) %>% 
  arrange(desc(Average_Checkouts))
```


# Programs & Outreach


## Data cleaning for P&G

```{r}
#Reading in the excel file based on the input from branch
oandp <- read_xlsx("Data/Example Data.xlsx", sheet = "outreach")

#Creating a synthetic month variable 
oandp$date <- as.Date(paste(oandp$year, oandp$month, "01", sep = "-"))

#grabbing the month and year (12-18 for december 2018) of the tweet tweeted for later analysis to group by month and year 
oandp$month <- format(oandp$date, "%b") 
```


```{r}
#Filtering specific dates we want to look at programs and outreach for 
#oandp <- oandp %>% filter(date <= as.Date("2023-03-01"))

#Figuring out attendance per programming event and outreach event
oandp <- oandp %>% group_by(month) %>% mutate(APO = attendence_o/total_outreach_events, APP = attendence_p/total_program_events)

#Rounding both new variables at once 
oandp <- oandp %>%
  mutate_at(vars(APO, APP), round, digits = 1)
```


## Summary and descriptive statistics for programming and outreach 
```{r}
#Quick summary table of attendance per outreach event and attendance per programming 
oandp %>% summarise(month, APP, APO)

oandp %>% filter(date >= as.Date("2023-04-01", "%Y-%m-%d")) #& date <= as.Date("2022-06-01", "%Y-%m-%d"))

#Looking at some summary statistics for the particular time (I think I'm doing pseudo math here)
summary(oandp$APO, na.rm = TRUE) #Be wary of this because of missing data 
summary(oandp$APP, na.rm = TRUE)
```


```{r eval=FALSE, include=FALSE}
#MISLEADING SUMMARY STATISTICS BECAUSE WHEN WE CALCULATE AN AVERAGE OF AN AVERAGE WE'RE NOT TAKING INTO ACCOUNT THE NUMBER OF PROGRAMS THAT WENT IN TO MAKING THAT AVERAGE IN THE FIRST PLACE 

#Quick summary table of attendance per outreach event and attendance per programming 
oandp %>% summarise(month, APP, APO)

summary(oandp$APO, na.rm = TRUE)
summary(oandp$APP, na.rm = TRUE)

testers <- oandp %>% filter(date >= as.Date("2022-01-01", "%Y-%m-%d") & date < as.Date("2022-04-01", "%Y-%m-%d"))

data.frame(unclass(summary(testers$APP)))
data.frame(unclass(summary(testers$APO)))
```


## Line graphs over time  
```{r}
#Attendance per outreach 

ggplot(oandp, aes(date, APO)) +
  geom_line(color = "blue3", size = 1) +
  geom_point(aes(size = total_outreach_events), alpha = 0.3, color = "black") +  
  geom_text(aes(label = total_outreach_events), size = 3, hjust = 0, vjust = -1, color = "black") +
  
  labs(title = "Attendence per Outreach Event at Library 1",
       x = "",
       y = "Attendence per Outreach",
       caption = "Figure 4.1",
       size = "Number of Outreach Events") +
  
    scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2022-01-01", "2023-06-01"), "%Y-%m-%d"),  
    date_labels = "%b '%y") +
  
    #scale_y_continuous(
      #breaks = seq(0, 1000, by = 100),
      #limits = c(0, 1000)) +
  
  scale_size_continuous(breaks = c(2, 4, 6), 
                        range = c(1, 8)) + # set the range of sizes here

  #scale_color_manual(name = NULL,
                     #values = c("APO" = "blue3", "APP" = "black")) +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = c(.76, .85),
        legend.direction = "horizontal",
        legend.title = element_text(size = 9), 
        legend.text = element_text(size = 8))
```


```{r}
#Attendance per program

ggplot(oandp, aes(date, APP)) +
  geom_line(color = "blue3", size = 1) +
  geom_point(aes(size = total_program_events), alpha = 0.3, color = "black") +  
  geom_text(aes(label = total_program_events), size = 3, hjust = 0, vjust = -0.8, color = "black") +
  
  labs(title = "Attendence per Program at Library 2",
       x = "",
       y = "Attendence per Program",
       caption = "Figure 4.2",
       size = "Number of Programming Events" ) +
  
    scale_x_date(
      date_breaks = "1 month",
      limits = as.Date(c("2022-01-01", "2023-06-01"), "%Y-%m-%d"),  
      date_labels = "%b '%y") +
  
    #scale_y_continuous(
      #breaks = seq(0, 70, by = 10),
      #limits = c(0, 70)) +
  
  scale_size_continuous(breaks = c(10, 30, 50), 
                        range = c(1, 8)) + # set the range of sizes here

  #scale_color_manual(name = NULL,
                     #values = c("APO" = "blue3", "APP" = "black")) +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = c(.34, .85),
        legend.direction = "horizontal",
        legend.title = element_text(size = 9), 
        legend.text = element_text(size = 8))
```


```{r}
#attendance per program year over year

ggplot(oandp, aes(month(date, label=TRUE, abbr=TRUE), APP, group = factor(year(date)), color = factor(year(date)))) +
  geom_line(size = 1) +
  geom_point(aes(size = total_program_events), alpha = 0.3, show.legend = FALSE) +  
  geom_text(aes(label = total_program_events), size = 3, hjust = 0, vjust = -0.8, show.legend = FALSE) +
  
  labs(title = "Attendence per Program at Library 2",
       x = NULL,
       y = "Attendence per Program",
       caption = "Figure 4.2",
       size = "Number of Programming Events", 
       subtitle = "*Size of Dot Represents Number of Programming Events") +
  
    #scale_x_date(
     #date_breaks = "1 month",
      #limits = as.Date(c("2022-01-01", "2023-06-01"), "%Y-%m-%d"),  
      #date_labels = " %b '%y") +
  
    scale_y_continuous(
      breaks = seq(0, 100, by = 10),
      limits = c(0, 100)) +
  
  scale_size_continuous(breaks = c(10, 30, 50), 
                        range = c(1, 8)) + # set the range of sizes here
  
  scale_color_manual(name = NULL,
                     values = c("2021" = "darkgreen", "2022" = "orangered2", "2023" = "blue3")) +

  theme_clean() +
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "grey", size = 0.2),
        panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
        legend.position = c(.84, .8),
        legend.text = element_text(size = 10))
```


```{r}
#attendance per outreach year over year

ggplot(oandp, aes(month(date, label=TRUE, abbr=TRUE), APO, group = factor(year(date)), color = factor(year(date)))) +
  geom_line(size = 1) +
  geom_point(aes(size = total_outreach_events), alpha = 0.3, show.legend = FALSE) +  
  geom_text(aes(label = total_outreach_events), size = 3, hjust = 0, vjust = -0.8, show.legend = FALSE) +
  
  labs(title = "Attendence per Outreach at Library 1",
       x = NULL,
       y = "Attendence per Outreach",
       caption = "Figure 4.2",
       size = "Number of Programming Events", 
       subtitle = "*Size of Dot Represents Number of Outreach Events") +
  
  #scale_x_date(
      #date_breaks = "1 month",
      #limits = as.Date(c("2022-01-01", "2023-06-01"), "%Y-%m-%d"),  
      #date_labels = " %b '%y") +
  
  #scale_y_continuous(
      #breaks = seq(0, 100, by = 10),
      #limits = c(0, 100)) +
  
  scale_size_continuous(breaks = c(10, 30, 50), 
                        range = c(1, 8)) + # set the range of sizes here
  
  scale_color_manual(name = NULL,
                     values = c("2021" = "darkgreen", "2022" = "orangered2", "2023" = "blue3")) +

  theme_clean() +
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "grey", size = 0.2),
        panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
        legend.position = c(.88, .8),
        legend.text = element_text(size = 10))
```




# Visitor Counts 


## Data wrangling for daily visitor counts

```{r}
visit_daily <- read_xlsx("Data/Example Data.xlsx", sheet = "daily_visits")

#modifying some variables 
visit_daily$day <- factor(visit_daily$day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

#list of branches to choose from 
unique(visit_daily$branch)

#Choosing a particular branch for analysis
branch_name <- "lib3"

visit_daily_bs <- filter(visit_daily, branch == branch_name)

#filtering the time we want to looks at (Q1 2023)
start_date_box <- "2023-01-01"
end_date_box <- "2023-9-11"
visit_daily_bst <- visit_daily_bs %>% filter(date >= as.Date(start_date_box, "%Y-%m-%d") & date <= as.Date(end_date_box, "%Y-%m-%d"))
```


### Daily boxplot for visitor count
```{r}
ggplot(visit_daily_bst, aes(day, door_count)) +
  
  geom_boxplot(fill = "blue1",  color = "black", outlier.color = "red") +
  
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2, fill = "white") + #adding the mean to the boxplots
  
  labs(title = paste0("Vistors Per Day at ",  branch_name),
       subtitle = paste0("From ", format(as.Date(start_date_box), "%B %d, %Y"), " to ", format(as.Date(end_date_box), "%B %d, %Y")),
       x = NULL,
       y = "Visitors",
       caption = "Figure 1.3") + 
  
  #scale_y_continuous(
    #breaks = seq(0, 700, by = 100),
    #limits = c(0, 700)) +
  
  #If we want to abbreviate dates
  #scale_x_discrete(labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
  
  #scale_fill_manual(values = c("grey85", "#FFDDCC"), guide = FALSE) +  #could potentially be used to highlight one of the boxplots 
  
  theme_clean() +
  theme(panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
        plot.subtitle = element_text(size = 10),
        plot.caption = element_text(size = 8))
```
### Summary statistics for daily visitor counts 
```{r}
summary(visit_daily_bst$door_count)

#Boxplot numbers but in chart form!

visit_daily_bst %>% 
  group_by(day) %>% 
  summarise(Average_Visitors = mean(door_count, na.rm = TRUE), n = n()) %>% 
  arrange(desc(Average_Visitors))

```



## Data wrangling for weekly visitor counts

```{r}
visitors <- read_xlsx("Data/Example data.xlsx", sheet = "visitor_count")

# First we need create a date column based on all the section headers so we need to reshape the data frame to long format 
visitors <- pivot_longer(visitors, cols = -Branch, names_to = "Week", values_to = "Count")

#With each variable we want now having it's own separate column, we can now pivot to wider data frame with each branch as a column header
visitors <- pivot_wider(visitors, names_from = Branch, values_from = Count)

#Tried but didn't work because of multiple months being in the same string 
#visitors$week_date <- sapply(visitors$Week, function(x) str_split(x, pattern = "-")[[1]][1])

visitors$w1 <- sapply(visitors$Week, function(x) trimws(str_split(x, pattern = "-")[[1]][1]))
visitors$w1 <- gsub("\\.", "", visitors$w1)

visitors$year <- sapply(visitors$Week, function(x) trimws(str_split(x, pattern = ",")[[1]][2]))

visitors$date_week <- paste0(visitors$w1, ", " , visitors$year)

visitors$date <- as.Date(visitors$date_week, format = "%b %d, %Y")

#Creating a quarters column based on the date so we can group the data easily by quarterly analysis 
visitors$quarter <- quarters(visitors$date)

#Creating a week day column so we can compare a weeks worth of data to the same week the year before
visitors$week_num <- week(visitors$date)
```


### visitor count by week for specific quarter year over year for one branch graph
```{r}
# Choosing branch for weekly visitor analysis

#Libraries to choose: "lib1" "lib2" "lib3" "lib4"

branch_weekly_visitors <- "lib1"
```


### Visitor count by week for large branches graph
```{r}
#we want to plot a line for each branch, so we need to specify the group aesthetic to ensure that ggplot knows to group the data by branch, rather than trying to connect points across different branches.

ggplot(visitors, aes(date)) +
  geom_line(aes(y = lib2, color = "lib2"), size = 1) +
  geom_line(aes(y = lib3, color = "lib3"), size = 1) +
  geom_line(aes(y = lib4, color = "lib4"), size = 1) +
  geom_line(aes(y = lib5, color = "lib5"), size = 1) +
  
  geom_line(aes(y = lib1, color = "lib1"),   size = 1) +
  geom_point(aes(y = lib1)) +

  labs(title = "Visitors Per Week Across all Branches",
       subtitle = "Q2 2023", 
       x= NULL,
       y = "Visitors",
       caption = "Figure 1.1") +
  
  scale_y_continuous(
    breaks = seq(0, 7000, by = 1000),
    limits = c(0, 7000)) +
  
  scale_x_date(
    date_breaks = "1 week",
    #Setting the dates here is kinda weird but this works! 
    limits = as.Date(c("2023-04-01", "2023-06-30"), "%Y-%m-%d"),  
    date_labels = "%B %d",
    expand = c(0.02, 0)) +
  
  scale_color_manual(name = NULL,
                     values = c("lib1" = "blue3", "lib2" = "black", "lib3" = "green", 
                               "lib4" = "red", "lib5" = "hotpink")) +
  
  
  theme_clean() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1),
        plot.subtitle = element_text(size = 12),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.72, 0.8),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8))
```



## Data wrangling for monthly visitor counts
```{r}
#Data wrangling for monthly visitor dataframe 
visitor_mon <- read_xlsx("Data/Example Data.xlsx", sheet = "visitor_mon")

#pivot to wider graph
visitor_mon <- pivot_wider(visitor_mon, names_from = branch, values_from = visitor_count)

#Creating a synthetic month variable 
visitor_mon$date <- as.Date(paste(visitor_mon$year, visitor_mon$month, "01", sep = "-"))

class(visitor_mon$date)

#grabbing the month and year (12-18 for december 2018) of the tweet tweeted for later analysis to group by month and year 
visitor_mon$month <- format(visitor_mon$date, "%b")
```


```{r}
# Choosing branch for monthly visitor analysis

#Libraries to choose: "lib1" "lib2" "lib3" "lib4"

branch_monthly_visitors <- "lib4"
```


### visitors per month year over year for 1 branch
```{r}
ggplot(visitor_mon, aes(month(date, label=TRUE, abbr=TRUE), .data[[branch_monthly_visitors]], group = factor(year), color = factor(year))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  
  
  labs(title = paste0("Visitors per Month at ", branch_monthly_visitors , " Year Over Year"),
       x= NULL,
       y = "Visitors",
       caption = "Figure 1.2",
       subtitle = "*Red highlight represents Q2",
       color = NULL) +
  
  #scale_y_continuous(
    #breaks = seq(0, 12000, by = 2000),
    #limits = c(0, 12000)) +
  
  scale_color_manual(name = NULL,
                     values = c("2021" = "chartreuse3", "2022" = "orangered2", "2023" = "blue3")) +

  annotate("rect", xmin = "Apr", xmax = "Jun", 
           ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  #annotate("rect", xmin = as.Date("2022-06-01", "%Y-%m-%d"), xmax = as.Date("2022-08-01", "%Y-%m-%d"), 
           #ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  theme_clean() +
  theme(#Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      plot.subtitle = element_text(size = 10),
      legend.position = c(0.85, 0.83)) +
    
    guides(fill = guide_legend(reverse = TRUE)) # Can't reverse legend order like this because year is a factor variable 
```


## Summary & descriptive statistics for Visitor count 
```{r}
#grouping visitor count by week to get day numbers that staff are familiar with
visitors %>% filter(date >= as.Date("2023-01-01", "%Y-%m-%d") & date < as.Date("2023-04-01", "%Y-%m-%d")) %>%
  group_by(date) %>% summarise(avg_per_day = lib1/7)

#Grabbing summary statistics for the lib2 branch over Q2 2023
vis <- visitors %>% filter(date >= as.Date("2023-04-01", "%Y-%m-%d") & date < as.Date("2023-06-30", "%Y-%m-%d"))

data.frame(unclass(summary(vis$lib2)))
```


# Circ per visitor  

## Data wrangling: modifying the visitor_mon dataframe in preperation for the merge with the circ_mon dataframe

```{r}
#Creating temporary data frames to remove date to then add the prefix to every other column name 
visitor_mon_nodate <- visitor_mon %>% select(-date)
visitor_mon_date <- visitor_mon %>% select(date)

#adding visitor to all colnames in that dataframe to help differentiate during merge
colnames(visitor_mon_nodate) <- paste("visitor_", colnames(visitor_mon_nodate), sep = "")

#Recombining the prefixed "visitor_" columns the date column dataframe
visitor_mon <- cbind(visitor_mon_nodate, visitor_mon_date)

#Now that the visitor_mon is ready we can merge it with the circ_mon 
combined_vist_circ <- merge(visitor_mon, circ_mon, by = "date")

#Sorting the combined_vist_circ dataframe to make looping through that dataframe go by date 
combined_vist_circ <- arrange(combined_vist_circ, by= "date")

#Creating 2 temporary datasets to get the branch names from the visitor_mon and circ_mon dataframe and save them to character vectors for the upcoming loop
beep <- visitor_mon %>% select(-visitor_month, -date, -visitor_year)
branches_vis <- colnames(beep)

beep <- circ_mon %>% select(-month, -date, -year)
branches_circ <- colnames(beep)


#### Making the merge of visitor_mon dataframe with the circ_mon dataframe #### 

#Creating an empty list to store our results
results <- list()

# loop over each restaurant
for (branch in branches_circ) {
  
  # calculate orders per visitor for the current restaurant
  circ_visitor <- combined_vist_circ[[branch]] / combined_vist_circ[[paste0("visitor_", branch)]] 
  
  # store the results in a new column in the data frame
  results[[paste0("circ_per_visitor", branch)]] <- circ_visitor
}

results <- as.data.frame(results)

circ_per_visitor <- cbind(combined_vist_circ$date, results)

#Renaming the date variable in the circ_per_visitor dataframe
circ_per_visitor <- circ_per_visitor %>%
  rename(date = "combined_vist_circ$date")

#Cleaning the temporary datasets 
rm(visitor_mon_nodate, visitor_mon_date, beep, circ_visitor, results)
```


## checkout per visitor for All branches
```{r}
#Look at poor Kellogsville :(

ggplot(circ_per_visitor, aes(date)) +
  geom_line(aes(y = circ_per_visitorlib1, color = "lib1"), size = 1) +
  geom_line(aes(y = circ_per_visitorlib2, color = "lib2"), size = 1) +  
  geom_line(aes(y = circ_per_visitorlib3, color = "lib3"), size = 1) +
  geom_line(aes(y = circ_per_visitorlib4, color = "lib4"), size = 1) +

  
  labs(
    title = "Checkouts per Visitor by Month",
    y = "Checkouts per Visitor",
    x = NULL,
    caption = "Figure 3.1",
    subtitle = "Large Branches") + 
  
  scale_y_continuous(
    breaks = seq(0, 15, by = 2),
    limits = c(0,15)) +
  
   scale_x_date(
    date_breaks = "1 month",
    limits = as.Date(c("2022-01-01", "2023-06-01"), "%Y-%m-%d"),
    date_labels = "%b '%y",
    expand = c(0.01, 0)) +
  
  scale_color_manual(name = NULL,
                     values = c("lib1" = "blue3", "lib2" = "black", "lib3" = "green", 
                                "lib4" = "red")) +
  
  
  #annotate("rect", xmin = as.Date("2022-06-01", "%Y-%m-%d"), xmax = as.Date("2022-08-01", "%Y-%m-%d"), 
           #ymin = -Inf, ymax = Inf, alpha = .1, fill = "red") +
  
  theme_clean() +
  
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        plot.subtitle = element_text(size = 10),
      #Adding a vertical line at each week interval 
      panel.grid.major.x = element_line(color = "grey", size = 0.2),
      panel.grid.minor.y = element_line(color = "grey", linetype = "dotted"),
      legend.position = c(0.73, 0.86),
        legend.direction = "horizontal",
        legend.text = element_text(size = 8)) 
```


# Statistical Analysis b/w checkouts and visitors


## Wrangling and merging the daily visitor and checkout dataframes together
```{r}
head(orig_circ_day)
head(visit_daily)

visit_daily_oncount <- select(visit_daily, date, day, branch, door_count)
#needed to properly merge on the date column in the orig_circ_day dataset 
visit_daily_oncount$date <- as.Date(visit_daily_oncount$date)

circday_visitday <- merge(orig_circ_day, visit_daily_oncount, by = c("date", "branch"))

summary(circday_visitday$transactions)
summary(circday_visitday$door_count)

#Assigning specific branch sizes
circday_visitday$branch_size <- ifelse(circday_visitday$branch %in% l_branches, "l_branch",
                                       ifelse(circday_visitday$branch %in% m_branches, "m_branch", "s_branch"))

#Filtering out small branches 
#circday_visitday <- circday_visitday %>% filter(branch_size == "l_branch")


#Dropping specific observations with outleirs based on their index number in the dataframe
#circday_visitday <- circday_visitday[-c(10630, 4588, 8191), ]
circday_visitday <- circday_visitday %>% filter(door_count <= 2000)

#filtering out the weekend
#circday_visitday <- circday_visitday %>% filter(!day.x %in% c("Saturday", "Sunday"))

```


```{r}
ggplot(circday_visitday, aes(door_count, transactions, color = branch_size)) +
  
  geom_point() +
  
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add a linear regression line
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +  # Add y = x line
  
       # Calculate and display Pearson's correlation coefficient
  geom_text(x = max(circday_visitday$door_count) * 0.10, 
            y = max(circday_visitday$transactions) * 0.65, 
            label = paste("Pearson's R =", round(cor(circday_visitday$door_count, circday_visitday$transactions), 2)),
            color = "hotpink") +
  
  labs(
      title = "Scatterplot Between Visitors & Checkouts",
      y = "Checkouts",
      x = "Visitors",
      caption = "Figure 5.1",
      color = NULL) + 
    
  scale_color_manual(values = c("l_branch" = "#f0645d", "m_branch" = "#fcb813", "s_branch" = "#00b059"), ##1e196a
                      labels = c("l_branch" = "Large Branch", "m_branch" = "Medium Branch", "s_branch" = "Small Branch")) +

  theme_clean() +
  
  theme(legend.position = c(0.13, 0.86),
        legend.text = element_text(size = 8)) 
```